Power and Pitfalls of Generic Smart Contracts

### 2022-02-03 cluster_computing/001_abstract.tex
cluster computing package

Generics are a powerful feature of programming languages that allows one
to write highly reusable code.
%
More specifically, they are based on the use of type placeholders in order
to produce parametrized code, that can be instantiated for each
concrete type provided for them. 
%
In many programming languages, such as Java, they are implemented by
\emph{erasure}\ie replaced by their upper bound type during compilation into bytecode.
%
This paper investigates the use of generics in the
context of smart contracts for blockchain, in order to implement
a contract for \emph{shared entities} (such as a company shared
by its shareholders),
by using the Hotmoka blockchain whose contracts are written in Java.
%
The considered case study is particularly important since
the validators' set of the blockchain itself is
a special case of shared entities.
The analysis shows that the power of generics comes at the risk of a too permissive
typing of the compiled code, due to the erasure mechanism, with a consequent possible attack
to the validators' set. This paper proposes a solution
that forces the compiler to generate more precise type information than
those arising by erasure.

>>>>

Generics are a powerful feature of programming languages that allows one
to write highly reusable code.
%
More specifically, they are based on the use of type placeholders in order
to produce parametrized code, that can be instantiated for each
concrete type provided for them. 
%
In many programming languages, such as Java, they are implemented by
\emph{erasure}\ie replaced by their upper bound type during compilation into bytecode.
%
This paper originated from a real security issue that we found
while using generics for writing
smart contracts for blockchain, in order to implement
a contract for \emph{shared entities} (such as a company shared
by its shareholders),
for the Hotmoka blockchain, whose contracts are written in Java.
%
The considered case study is particularly important since
the validators' set of the blockchain itself is
a special case of shared entities.
The analysis shows that the power of generics comes at the risk of a too permissive
typing of the compiled code, due to the erasure mechanism, with a consequent possible attack
to the validators' set. This paper proposes a solution
that forces the compiler to generate more precise type information than
those arising by erasure.

### 2022-02-03 cluster_computing/010_introduction.tex
cluster computing package

\IEEEPARstart{B}{lockchains} exploit the redundant, concurrent execution of the same
transactions on a decentralized network of many machines,
in order to enforce their execution in accordance with
a set of predefined rules. Namely, blockchains make it hard, for a single machine,
to disrupt the semantics of the transactions or their ordering: a misbehaving single machine
gets immediately put out of consensus and isolated. Bitcoin~\cite{Nakamoto08,book-mastering-bitcoin}
has been the first blockchain's success story. Here
transactions are programmed in a non-Turing complete bytecode language,
almost exclusively used to implement transfers of units of coins between \emph{accounts}.

>>>>

Blockchains exploit the redundant, concurrent execution of the same
transactions on a decentralized network of many machines,
in order to enforce their execution in accordance with
a set of predefined rules. Namely, blockchains make it hard, for a single machine,
to disrupt the semantics of the transactions or their ordering: a misbehaving single machine
gets immediately put out of consensus and isolated. Bitcoin~\cite{Nakamoto08,book-mastering-bitcoin}
has been the first blockchain's success story. Here
transactions are programmed in a non-Turing complete bytecode language,
almost exclusively used to implement transfers of units of coins between \emph{accounts}.

### 2022-02-03 cluster_computing/050_related_work.tex
cluster computing package

Finally, as regards to the definition of new programming languages for safe smart contracs,
Scilla~\cite{scilla} has been tailored by taking System~F as a foundational calculus.
It is able to provide strong safety guarantees by means of type soundness.
Thanks to its minimalistic nature, it has been possible to define also a generic and extensible
framework for lightweight verification of smart contracts by means of user-defined domain-specific
analyses. The type variables of the functional foundational calculus can be seen as
generic types. We do not know how they are compiled and if the strong typing guarantee of the
source code extends to the compiled code as well. Scilla contracts are developed with
the Neo Savant online IDE. Currently, neither Neo Savant IDE nor the block explorer
allow one to inspect the compiled bytecode, in order to undertsnad how generic types are compiled.

>>>>

Finally, as regards to the definition of new programming languages for safe smart contracs,
Scilla~\cite{scilla} has been tailored by taking System~F as a foundational calculus.
It is able to provide strong safety guarantees by means of type soundness.
Thanks to its minimalistic nature, it has been possible to define also a generic and extensible
framework for lightweight verification of smart contracts by means of user-defined domain-specific
analyses. The type variables of the functional foundational calculus can be seen as
generic types. We do not know how they are compiled and if the strong typing guarantee of the
source code extends to the compiled code as well. Scilla contracts are developed with
the Neo Savant online IDE. Currently, neither Neo Savant IDE nor the block explorer
allow one to inspect the compiled bytecode, in order to understand how generic types are compiled.
